FEM Model Analysis

### 2021-06-28 FEM_R/model_analysis_core.Rmd
Fixed crossvalidation script error, also reduced the analysis to bare minimum for speedy runs

#AgeUK
male_au <- read.csv('../../../Validation/AgeUK-chronDisease-MALE.csv')
female_au <- read.csv('../../../Validation/AgeUK-chronDisease-FEMALE.csv')

# FEM 2014 detailed output
#FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
## extract males and females
#mFEM_2014 <- FEM_2014 %>% filter(male == 1)
#fFEM_2014 <- FEM_2014 %>% filter(male == 0)

# FEM 2014 summary
#FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
FEM <- read_dta('ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')

# Keep only year 2014 from FEM summary
FEM <- filter(FEM, year == 2014)

# Specify the age groups to extract from FEM
ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
# Add m and f for male and female variables
m_ages <- paste0('m_', ages)
f_ages <- paste0('f_', ages)

# Collect the age suffixes into a vector
keep_vars <- c(m_ages, f_ages)
# Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')

# Don't like using loops but this one was just too simple to bother with figuring out the apply version
final <- c()
ticker <- 1
# Generating the column (variable) names we want to keep in a loop
for(item in diseases) {
  for(anotherItem in keep_vars) {
    final[ticker] <- paste0('p_', item, '_', anotherItem)
    ticker <- ticker + 1
  }
}

# Keep the right vars using varnames generated in loop
FEM <- FEM %>%
          select(c('year', all_of(final)))

# Rename some vars to remove 'p_' before merging
FEM <- FEM %>%
          rename_at(vars(starts_with('p_')),
                    funs(str_replace(., 'p_', ''))) %>%
          select(!year)

pivoted.FEM <- pivot_longer(data = FEM,
                            cols = hibpe_m_6064:asthmae_f_100p, 
                            names_to = c('Condition', 'Gender', 'Age_Group'), 
                            names_sep = '_')

# Convert value from fraction to percentage
pivoted.FEM$value <- pivoted.FEM$value * 100

# Rename cols to enable reshaping
male_au <- male_au %>%
                dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
                       m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
                       m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
                select(!Total)

female_au <- female_au %>%
                dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
                       f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
                       f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
                select(!Total)

# Now need to reshape these datasets to long format
pivoted.male <- pivot_longer(data = male_au,
                             cols = m_6064:m_100p,
                             names_to = c('Gender', 'Age_Group'),
                             names_sep = '_')

pivoted.female <- pivot_longer(data = female_au,
                             cols = f_6064:f_100p,
                             names_to = c('Gender', 'Age_Group'),
                             names_sep = '_')

# Now row bind these frames together, combining male and female data
ageUK <-rbind(pivoted.male, pivoted.female)

# Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
# extract all heart related conditions for combining
#ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
# Aggregate the values by gender and age group
#ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
# Rename columns and add condition column back
#colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
#ageUK_heart$Condition <- 'Heart Problems'

# Now drop the heart problems from original ageUK and replace with aggregate
ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
#ageUK <- rbind(ageUK, ageUK_heart)

# Rename a few just for completeness
#ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
#ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)

# Finally, add column to specify that this is ageUK data
ageUK$source <- "ageUK"

# REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
#ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')

# Make copy
newFEM <- pivoted.FEM

# Rename all the conditions to fit ageUK
# NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
#       Therefore only validation from this will be that FEM is higher than ageUK
#newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
#newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
#newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)

#force Make to run this again

# Add column to specify this is FEM data
newFEM$source <- "FEM"

# Now row bind the datasets
combined <- rbind(newFEM, ageUK)
# Specify factor levels for age group
combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
# Sort
combined <- arrange(combined, source, Condition, Gender, Age_Group)

# Function to filter the combined dataset and visualise
visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {

  if(gender == 'm') {
    full_gender <- 'Male'
  }
  else if(gender == 'f') {
    full_gender <- 'Female'
  }

  # filter by gender and condition
  data <- data %>%
            filter(Gender == gender) %>%
            filter(Condition == cd)

  # now plot!
  p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
          geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
          labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')

  # Save plot into in E-FEM_Paper/figures/
  ggsave(filename = paste0(cd, '_', gender, '.png'),
         plot = p,
         path = path,
         width = 6,
         height = 4)

  return(p)
}

@@

```{r}
# Collect all conditions (and both genders) into a list, and apply the function above to that list
cd_list <- unique(combined$Condition)
gender_list <- unique(combined$Gender)

for(cd in cd_list) {
  for(gen in gender_list) {
    plot <- visAgeUKValid(combined, cd, gen)
    print(plot)
  }
}
```

>>>>

# #AgeUK
# male_au <- read.csv('../../../Validation/AgeUK-chronDisease-MALE.csv')
# female_au <- read.csv('../../../Validation/AgeUK-chronDisease-FEMALE.csv')
# 
# # FEM 2014 detailed output
# #FEM_2014 <- read_dta('ELSA_Baseline/detailed_output/y2014_rep1.dta')
# ## extract males and females
# #mFEM_2014 <- FEM_2014 %>% filter(male == 1)
# #fFEM_2014 <- FEM_2014 %>% filter(male == 0)
# 
# # FEM 2014 summary
# #FEM <- read_dta('ELSA_Baseline/ELSA_Baseline_summary.dta')
# FEM <- read_dta('ELSA_AgeUK_core_valid/ELSA_AgeUK_core_valid_summary.dta')
# 
# # Keep only year 2014 from FEM summary
# FEM <- filter(FEM, year == 2014)
# 
# # Specify the age groups to extract from FEM
# ages <- c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p')
# # Add m and f for male and female variables
# m_ages <- paste0('m_', ages)
# f_ages <- paste0('f_', ages)
# 
# # Collect the age suffixes into a vector
# keep_vars <- c(m_ages, f_ages)
# # Now set up the chronic diseases in a list so we can create a bigger list with all of them produced
# diseases <- c('hibpe', 'stroke', 'diabe', 'demene', 'asthmae')
# 
# # Don't like using loops but this one was just too simple to bother with figuring out the apply version
# final <- c()
# ticker <- 1
# # Generating the column (variable) names we want to keep in a loop
# for(item in diseases) {
#   for(anotherItem in keep_vars) {
#     final[ticker] <- paste0('p_', item, '_', anotherItem)
#     ticker <- ticker + 1
#   }
# }
# 
# # Keep the right vars using varnames generated in loop
# FEM <- FEM %>%
#           select(c('year', all_of(final)))
# 
# # Rename some vars to remove 'p_' before merging
# FEM <- FEM %>%
#           rename_at(vars(starts_with('p_')),
#                     funs(str_replace(., 'p_', ''))) %>%
#           select(!year)
# 
# pivoted.FEM <- pivot_longer(data = FEM,
#                             cols = hibpe_m_6064:asthmae_f_100p, 
#                             names_to = c('Condition', 'Gender', 'Age_Group'), 
#                             names_sep = '_')
# 
# # Convert value from fraction to percentage
# pivoted.FEM$value <- pivoted.FEM$value * 100
# 
# # Rename cols to enable reshaping
# male_au <- male_au %>%
#                 dplyr::rename(m_6064 = X60.64, m_6569 = X65.69, m_7074 = X70.74, 
#                        m_7579 = X75.79, m_8084 = X80.84, m_8589 = X85.89, 
#                        m_9094 = X90.94, m_9599 = X95.99, m_100p = Over.100) %>%
#                 select(!Total)
# 
# female_au <- female_au %>%
#                 dplyr::rename(f_6064 = X60.64, f_6569 = X65.69, f_7074 = X70.74, 
#                        f_7579 = X75.79, f_8084 = X80.84, f_8589 = X85.89, 
#                        f_9094 = X90.94, f_9599 = X95.99, f_100p = Over.100) %>%
#                 select(!Total)
# 
# # Now need to reshape these datasets to long format
# pivoted.male <- pivot_longer(data = male_au,
#                              cols = m_6064:m_100p,
#                              names_to = c('Gender', 'Age_Group'),
#                              names_sep = '_')
# 
# pivoted.female <- pivot_longer(data = female_au,
#                              cols = f_6064:f_100p,
#                              names_to = c('Gender', 'Age_Group'),
#                              names_sep = '_')
# 
# # Now row bind these frames together, combining male and female data
# ageUK <-rbind(pivoted.male, pivoted.female)
# 
# # Now need to deal with the fact that ageUK data splits heart issues into 3 different conditions
# # extract all heart related conditions for combining
# #ageUK_heart <- filter(ageUK, Condition == c('Coronary Heart Disease', 'Heart Failure', 'Atrial Fibrillation'))
# # Aggregate the values by gender and age group
# #ageUK_heart <- aggregate(ageUK_heart$value, by = list(ageUK_heart$Gender, ageUK_heart$Age_Group), FUN = sum)
# # Rename columns and add condition column back
# #colnames(ageUK_heart) <- c('Gender', 'Age_Group', 'value')
# #ageUK_heart$Condition <- 'Heart Problems'
# 
# # Now drop the heart problems from original ageUK and replace with aggregate
# ageUK <- filter(ageUK, Condition != 'Coronary Heart Disease', Condition != 'Heart Failure', Condition != 'Atrial Fibrillation')
# #ageUK <- rbind(ageUK, ageUK_heart)
# 
# # Rename a few just for completeness
# #ageUK$Condition <- gsub("Cancer, recent", "Cancer", ageUK$Condition)
# #ageUK$Condition <- gsub("COPD", "Lung Disease (COPD)", ageUK$Condition)
# 
# # Finally, add column to specify that this is ageUK data
# ageUK$source <- "ageUK"
# 
# # REMOVE CONDITIONS FROM ageUK THAT ARE NOT IN FEM (may change in future)
# #ageUK <- filter(ageUK, Condition != c('Asthma', 'Depression', 'Severe Mental Health Condition'))
# ageUK <- filter(ageUK, Condition != 'Depression', Condition != 'Severe Mental Health Conditions')
# ageUK <- filter(ageUK, Condition != 'Cancer, recent', Condition != 'COPD')
# 
# # Make copy
# newFEM <- pivoted.FEM
# 
# # Rename all the conditions to fit ageUK
# # NOTE: Lung disease is not a perfect match, as ageUK only contains info on COPD
# #       Therefore only validation from this will be that FEM is higher than ageUK
# #newFEM$Condition <- gsub("hearte", "Heart Problems", newFEM$Condition)
# newFEM$Condition <- gsub("hibpe", "Hypertension", newFEM$Condition)
# newFEM$Condition <- gsub("stroke", "Stroke", newFEM$Condition)
# newFEM$Condition <- gsub("diabe", "Diabetes", newFEM$Condition)
# #newFEM$Condition <- gsub("lunge", "Lung Disease (COPD)", newFEM$Condition)
# #newFEM$Condition <- gsub("cancre", "Cancer", newFEM$Condition)
# newFEM$Condition <- gsub("demene", "Dementia", newFEM$Condition)
# newFEM$Condition <- gsub("asthmae", "Asthma", newFEM$Condition)
# 
# 
# #force Make to run this again
# 
# # Add column to specify this is FEM data
# newFEM$source <- "FEM"
# 
# # Now row bind the datasets
# combined <- rbind(newFEM, ageUK)
# # Specify factor levels for age group
# combined$Age_Group <- factor(combined$Age_Group, levels = c('6064', '6569', '7074', '7579', '8084', '8589', '9094', '9599', '100p'))
# # Sort
# combined <- arrange(combined, source, Condition, Gender, Age_Group)
# 
# # Function to filter the combined dataset and visualise
# visAgeUKValid <- function(data, cd, gender, path = '/home/luke/Documents/E_FEM_clean/Deliverables/E-FEM_Paper/figures/validation/ageUK/') {
#   
#   if(gender == 'm') {
#     full_gender <- 'Male'
#   }
#   else if(gender == 'f') {
#     full_gender <- 'Female'
#   }
#   
#   # filter by gender and condition
#   data <- data %>%
#             filter(Gender == gender) %>%
#             filter(Condition == cd)
#   
#   # now plot!
#   p <- ggplot(data = data, aes(x = Age_Group, y = value)) +
#           geom_bar(aes(fill = source), stat = "identity", position = "dodge") +
#           labs(title = paste(cd, full_gender, sep = ' '), x = 'Age Group', y = 'Proportion')
#   
#   # Save plot into in E-FEM_Paper/figures/
#   ggsave(filename = paste0(cd, '_', gender, '.png'),
#          plot = p,
#          path = path,
#          width = 6,
#          height = 4)
#   
#   return(p)
# }

@@

```{r echo=FALSE}
# # Collect all conditions (and both genders) into a list, and apply the function above to that list
# cd_list <- unique(combined$Condition)
# gender_list <- unique(combined$Gender)
# 
# for(cd in cd_list) {
#   for(gen in gender_list) {
#     plot <- visAgeUKValid(combined, cd, gen)
#     print(plot)
#   }
# }
```
